Understanding the Lomb-Scargle Periodogram

### 2017-01-14 PracticalLombScargle.tex
Add some discussion of uncertainties

All figures in this work were produced using Python, and in particular the
Numpy \citep{numpy, numpybook},
Pandas \citep{pandas},
Astropy \citep{Astropy2013},
and Matplotlib \citep{matplotlib} packages.
Code to reproduce these figures is available in IPython notebooks at \url{http://github.com/jakevdp/PracticalLombScargle/}.

@@

While Nyquist arguments based on average or minimum sampling fail spectacularly,
there is a sense in which the Nyquist limit can be applied to unevenly-spaced
data. \citet{Eyer99} explore this issue in some detail, and in particular
proves the following:
\begin{quote}
Let $p$ be the largest value such that each $t_i$ can be written $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then is $f_{Ny} = 1 / (2p)$.
\end{quote}
In other words, computing the Nyquist rate for unevenly-spaced data requires
finding some common factor, such that each spacing $\Delta t_i$ is {\it exactly}
an integer multiple of this factor.
This can be proven formally, but can be understood by thinking of such data as
a windowed version of uniformly sampled data with spacing $p$, where the window
is zero at some observations.
Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window
function applied to that data does not change that fact.

@@

Partly due to the deep connection between Fourier analysis and least-squares
analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
has since become commonly referred to as the {\it Lomb-Scargle Periodogram}.

@@

\subsection{Observational Noise}
\label{extensions-observational-noise}
Perhaps the most important modification of the periodogram is to
construct a periodogram that correctly handles measurement error in the data.
This involves changing the expression for the $\chi^2$ statistic in a manner
analogous to the change for a regression problem.
If there are Gaussian errors $\sigma_n$ on each observed $y_n$, we can re-write \eq{chi2-simple} in the following (standard) way:
\begin{equation}
  \chi^2(f) \equiv \sum_n \left(\frac{y_n - y_{model}(t_n;f)}{\sigma_n}\right)^2
  \eqlabel{chi2-with-errors}
\end{equation}
The periodogram constructed from this $\chi^2$ definition will more accurately
reflect the spectral power of noisy observations.
The effect of this modification on \eq{lomb-scargle-periodogram} is the addition
of a multiplicative weight $1/\sigma_n$ within each of the summations.
Early versions of this sort of modification appeared in \citet{Gilliland87}
and \citet{Irwin89}; \citet{Zechmeister09} showed that such a modification
does not change the desirable statistical properties of the resulting
periodogram.

@@

Another commonly-applied modification of the periodogram has variously been
called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency\footnote{We choose to follow \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to avoid confusion surrounding the different meanings of the term ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and \citet{Zechmeister09}.}:
\begin{equation}
  y_{model}(t;f) = y_{0, f} + A_f \sin(2 \pi f (t - \phi_f))
  \eqlabel{floating-mean-model}
\end{equation}

This is important, because the standard Lomb-Scargle approach assumes that the
data are pre-centered around the mean value of the signal.
In the past this was primarily accomplished by pre-centering the data before
the Lomb-Scargle fit using the mean estimated from the data points themselves.

With full phase coverage of the observed signal, this approach is generally
robust. However, due to selection effects and survey cadence, full phase
coverage can not always be guaranteed, which can potentially lead to situations
where the periodogram entirely misses peaks of interest.
\Fig{standard-vs-floatingmean} shows a simulated example of this: the data
consist of noisy observations of a sinusoidal signal in which the faintest
observations are omitted from the dataset (due to, {e.g.} a detection threshold).
Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
periodogram that misses the true period of 0.3 days (lower right panel).
Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
that detects this true period (upper right panel).

@@

A further generalization involves multi-term Fourier models at each frequency:
rather than fitting just a single sinusoid at each frequency, we might fit a
partial Fourier series at each frequency, adding $K$ additional components at
integer multiples of the fundamental frequency:
\begin{equation}
  y_{model}(t;f) = y^{0}_f + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
\end{equation}
\citet{Bretthorst88} takes a comprehensive look at this type of multi-term
extension to the periodogram, as well as extensions to decaying signals and
other more complex models.

@@

\subsection{A Note About ``Bayesian'' Periodograms}
\sectlabel{bayesian-periodograms}
The least squares view of the Lomb-Scargle periodogram creates a natural
bridge, via maximum likelihood, to Bayesian approaches to periodograms.
In fact, \citet{Jaynes87} showed that the standard form of the Lomb-Scargle
periodogram can be derived directly from the principles of Bayesian Probability
Theory outlined in his comprehensive treatment of the subject \citep{Jaynes03}.
In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
for detecting a stationary sinusoidal signal in the presence of Gaussian
noise.

For the standard model, the Bayesian periodogram is given by
\begin{equation}
  p(f~|~D) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
  \eqlabel{bayesian-periodogram}
\end{equation}
where $P_{LS}(f)$ is the Lomb-Scargle power from \eq{lomb-scargle-periodogram}.
The effect of this exponentiation, as seen in \fig{bayesian-periodogram},
is to suppress all peaks in relation to the largest one of the spectrum.

@@

The problem, however, is that the Bayesian periodogram does {\it not} return
a simple probability of the data showing a particular frequency, but rather
a probability of each frequency conditioned on a relatively strong assumption:
that the data are drawn from a sinusoidal model.
Now, the standard Lomb-Scargle periodogram also assumes a sinusoidal model,
and is thus subject to the same criticism.
But unlike the simple Bayesian approach, the standard periodogram affords
interpretation in light of Fourier transforms, window functions, and the
like: in most cases of interest, this turns out to be {\it vital} to the
application and interpretation of the periodogram in most cases of interest
(see \sect{failure-modes}).

In other words, the Bayesian approach goes ``all-in'' on the least squares
interpretation of the periodogram, exponentially suppressing the information
that allows you to reason about the periodogram in light of its connection
to Fourier analysis.
In contrived cases with structureless window functions, exponential attenuation
of side-peaks and side-lobes may seem appealing \citep[see, e.g.][]{Mortier15},
but that appeal extends to the real world only in special cases.
That is not to say that all Bayesian approaches to periodic analysis are
similarly flawed; there have been many interesting studies that go beyond
the simple sinusoidal model and use more complex and/or flexible models.
Some examples are models based on Fourier
extensions with strong priors \citep[e.g.][]{Bretthorst88}, flexible
non-parametric functions \citep[e.g.][]{Gregory92}, constrained stochastic
models \citep[e.g.][]{Kelly14}, and Gaussian process models \citep{Angus16}.
Though these can often be more accurate and powerful than the Lomb-Scargle
approach, they tend to be far more expensive computationally.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig25_failure_modes}
  \caption{Comparison of the true period and peak Lomb-Scargle period for
    1000 simulated periodic light curves.
    Each has 60 irregular observations over 180 nights,
    with a cadence typical of ground-based surveys ({i.e.} showing a strong
    diurnal window pattern similar to \fig{LINEAR-window}).
    The Lomb-Scargle peak does not always coincide with the true period,
    and there is noticeable structure among these failures.
    Panels (a)-(d) show some of the failure modes we can expect due to this
    window function; see the text for more discussion.
    \figlabel{failure-modes}}
\end{figure}

@@

These patterns can be understood in terms of the interaction between the
window function and the underlying spectral power.
As we discussed in \sect{window-functions},
a nightly observation pattern---typical of ground-based surveys---leads to
a window function with a strong diurnal component that causes each frequency
signature $f_0$ to be partially aliased at $f_0 + n \delta f$, for integers
$n$ and $\delta f = 1\ {\rm day}^{-1}$
(Recall \figs{LINEAR-window}{LINEAR-window-effect}).
This is the source of the failure modes depicted in panel (a) of
\fig{failure-modes}; the periods shown are
\begin{equation}
  P_{obs} = \left(\frac{1}{P_{true}} + n\delta f\right)^{-1}
\end{equation}
and for this data, close to 36\% of the objects are mis-characterized this way.

Another mode of failure is when the periodogram for an object with frequency
$f_0$ isolates a higher harmonic of that fundamental frequency, such as $2f_0$.
This can occur for signals that are not strictly sinusoidal, and this harmonic
frequency peak is subject to the same windowing effects as above.
Thus we can describe the failure modes depicted in panel (b) of \fig{failure-modes} with the following equation, for integers $m > 0$:
\begin{equation}
  P_{obs} = \left(\frac{m}{P_{true}} + n\delta f\right)^{-1}
\end{equation}
In \fig{failure-modes}, the $m=2$ harmonic and its aliases account for roughly
15\% of the results.

Panel (c) of \fig{failure-modes} shows the final pattern of failure in
Lomb-Scarlge results.
This is closely related to panels (a) and (b), but comes from the fact that
the Lomb-Scargle periodogram at frequency $f_0$ contains contributions from
signals at both $+f_0$ and $-f_0$.
When diurnal aliasing leads to power at negative frequencies, this is reflected
at the absolute value of that frequency, so it is more accurate to take the
absolute value of the above formula
\begin{equation}
  P_{obs} = \left|\frac{m}{P_{true}} + n\delta f\right|^{-1}
  \eqlabel{failure-modes}
\end{equation}
Panel (c) shows the 5\% of objects that fall along this failure mode for
$m=1$, $n=-2$.
After accounting for all these known sources of failure, only roughly
1\% of points are misclassified in an ``unexplained'' way, as seen in panel (d).

When applying a periodogram in practice, it is vital to take such effects into
account, rather than blindly relying on the single periodogram peak as your
best estimate of the period.
Applying understanding of windowing and aliasing effects can help in detecting
failures of the periodogram, but is no silver bullet.
For an observed peak at $f_{peak}$ from a survey whose window has strong power
at $\delta f$, something like the following should probably be employed:
\begin{enumerate}
  \item Check for a peak at $f_{peak}/m$ for at least $m \in \{2, 3\}$.
    If a significant peak is found, then $f_{peak}$ is probably an order-$m$
    harmonic of the true frequency.
  \item Check for peaks at $|f_{peak} \pm n\delta f|$ for at least
    $n \in \{1, 2\}$. If these aliases exist, then it is likely that you have
    found a peak on the sequence of expected aliases---though keep in mind that
    there is no way to know from the periodogram alone whether or not
    this is the ``true'' peak!
  \item For each of the top few of these aliases, fit a more complicated model
    (such as a multi-term Fourier series, template-based fit, etc.) to select
    among them.
\end{enumerate}
For noisy observations, this procedure cannot generally guarantee that you
have found the correct peak, but it is far preferable than making the mistake
of blindly trusting the highest peak in the periodogram!
We will come back to the question of uncertainty in the periodogram result
in \sect{uncertainties}

\subsection{Algorithmic Considerations}
\sectlabel{algorithmic-considerations}
Given the size of the frequency grid required to fully characterize the periods
from a given dataset, it is vital to have an efficient algorithm for evaluating
the periodogram. Unfortunately, the periodogram itself is a relatively slow
operation.

Suppose you have a set of $N$ observations over a time-span $T$ for an average
cadence of $\overline{\delta t} = N/T$.
From \eq{n-eval} we see that the number of frequencies evaluated is
$N_f \propto T f_{max} = N f_{max} / \overline{\delta t}$.
Holding constant the average survey cadence $\overline{\delta t}$ and individual
observation properties (which control $f_{max}$), we find that the number of
frequencies is directly proportional to the number of data points.
The computation of the Lomb-scargle periodogram in \eq{lomb-scargle-periodogram}
requires sums over $N$ sinusoids at each of the
$N_f$ frequencies, and thus we see that the
na{\"i}ve scaling of the algorithm with the size of the dataset is
$\mathcal{O}(N^2)$, when survey properties are held constant.
Due to the trigonometric functions involved, this turns out to be a rather
``expensive'' $\mathcal{O}(N^2)$, which makes the straightforward
implementation impractical for even modestly-sized datasets.

@@

All periodograms in this paper were computed using the Astropy
implementations of the \citet{Press89}, \citet{Zechmeister09}, and
\citet{Palmer09} algorithms, which were adapted from Python code originally
published by \citet{ICVG2014} and \citet{VanderPlas2015}.
All code for this paper is available in the form of Jupyter notebooks in the
paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}.

@@

\begin{itemize}
  \item $\chi^2$ normalization
  \item PSD normalization
  \item Other normalizations \citep{Baluev2008}
\end{itemize}

\section{Uncertainties in Periods}
\sectlabel{uncertainties}

\begin{itemize}
  \item Peak Width and why it's a poor proxy
  \item False Alarm Probability and its pitfalls
  \item Baluev method and why it's not enough
  \item Motivation and example of Bootstrap
\end{itemize}

>>>>

All figures in this work were produced using Python, and in particular the
Numpy \citep{numpy, numpybook},
Pandas \citep{pandas},
Astropy \citep{Astropy2013},
and Matplotlib \citep{matplotlib} packages.
Code to reproduce these figures is available in Jupyter notebooks at
\url{http://github.com/jakevdp/PracticalLombScargle/}.

@@

While Nyquist arguments based on average or minimum sampling fail spectacularly,
there is a sense in which the Nyquist limit can be applied to unevenly-spaced
data. \citet{Eyer99} explore this issue in some detail, and in particular
prove the following:
\begin{quote}
Let $p$ be the largest value such that each $t_i$ can be written $t_i = t_0 + n_i p$, for integers $n_i$. The Nyquist Frequency then is $f_{Ny} = 1 / (2p)$.
\end{quote}
In other words, computing the Nyquist rate for unevenly-spaced data requires
finding some common factor, such that each spacing $\Delta t_i$ is {\it exactly}
an integer multiple of this factor.
\citet{Eyer99} prove this formally, but the result can be understood
by thinking of such data as a windowed version of uniformly sampled
data with spacing $p$, where the window is zero at some observations.
Such uniform data has a classical Nyquist limit of $1/(2p)$, and a window
function applied to that data does not change that fact.

@@

Partly due to the deep connection between Fourier analysis and least-squares
analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
has since become commonly referred to as the {\it Lomb-Scargle Periodogram}
\citep[though versions of this approach had been employed even earlier;
  see, e.g.][]{Gottlieb75}.

@@

\subsection{Observational Noise}
\sectlabel{extensions-observational-noise}
Perhaps the most important modification of the periodogram is to
construct a periodogram that correctly handles measurement error in the data.
This involves changing the expression for the $\chi^2$ statistic in a manner
analogous to the change for a regression problem.
If there are Gaussian errors $\sigma_n$ on each observed $y_n$, we can re-write \eq{chi2-simple} in the following (standard) way:
\begin{equation}
  \chi^2(f) \equiv \sum_n \left(\frac{y_n - y_{model}(t_n;f)}{\sigma_n}\right)^2
  \eqlabel{chi2-with-errors}
\end{equation}
The periodogram constructed from this $\chi^2$ definition will more accurately
reflect the spectral power of noisy observations.
The effect of this modification on \eq{lomb-scargle-periodogram} is the addition
of a multiplicative weight $1/\sigma_n$ within each of the summations.
Early versions of this sort of modification appeared in \citet{Gilliland87}
and \citet{Irwin89}; \citet{Zechmeister09} showed that such a modification
does not change the desirable statistical properties of the resulting
periodogram.

@@

Another commonly-applied modification of the periodogram has variously been
called the {\it Date-compensated Discrete Fourier Transform} \citep{Ferraz-Mello81}, the {\it floating-mean periodogram} \citep{Cumming99}, and the {\it generalized Lomb-Scargle Method} \citep{Zechmeister09}, and involves adding an offset term to the sinusoidal model at each frequency\footnote{We choose to follow \citet{Cumming99} and \citet{VanderPlas2015} and call this a ``Floating Mean'' model, to avoid confusion surrounding the different uses of the term ``generalized periodogram'' in, {e.g.} \citet{Bretthorst88} and \citet{Zechmeister09}.}:
\begin{equation}
  y_{model}(t;f) = y_0(f) + A_f \sin(2 \pi f (t - \phi_f))
  \eqlabel{floating-mean-model}
\end{equation}
This turns out to be quite important in practice, because the standard
Lomb-Scargle approach assumes that the
data are pre-centered around the mean value of the (unknown) signal.
In many analyses, this requirement is accomplished by pre-centering using
the mean estimated from the data points themselves.

When the data provide full phase coverage of the observed signal,
this approach is generally robust.
However, due to selection effects and survey cadence, full phase
coverage can not always be guaranteed, which can potentially lead to situations
where the periodogram entirely misses peaks of interest.
\Fig{standard-vs-floatingmean} shows a simulated example of this: the data
consist of noisy observations of a sinusoidal signal in which the faintest
observations are omitted from the dataset (due to, {e.g.} a detection threshold).
Applying the standard Lomb-Scargle periodogram to pre-centered data leads to a
periodogram that misses the true period of 0.3 days (lower right panel).
Using the floating-mean model of \eq{floating-mean-model} yields a periodogram
that detects this true period (upper right panel).

@@

A further generalization involves multi-term Fourier models at each frequency:
rather than fitting just a single sinusoid at each frequency, we might fit a
partial Fourier series at each frequency, adding $K$ additional components at
integer multiples of the fundamental frequency:
\begin{equation}
  y_{model}(t;f) = A_f^{0} + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
\end{equation}
\citet{Bretthorst88} takes a comprehensive look at this type of multi-term
extension to the periodogram, as well as extensions to decaying signals and
other more complex models.

@@

\subsection{A Note About ``Bayesian'' Periodograms}
\sectlabel{bayesian-periodograms}
The least squares view of the Lomb-Scargle periodogram creates a natural
bridge, via maximum likelihood, to Bayesian periodic analysis.
In fact, \citet{Jaynes87} showed that the standard form of the Lomb-Scargle
periodogram can be derived directly from the principles of Bayesian probability
theory outlined in his comprehensive treatment of the subject \citep{Jaynes03}.
In the Bayesian view, the Lomb-Scargle periodogram is the optimal statistic
for detecting a stationary sinusoidal signal in the presence of Gaussian
noise.

For the standard model, the Bayesian periodogram is given by the posterior
probability of frequency $f$ given the data $D$ and sinusoidal model $M$:
\begin{equation}
  p(f~|~D, M) \propto e^{-N \chi^2 / 2} \propto e^{P_{LS}(f)}
  \eqlabel{bayesian-periodogram}
\end{equation}
where $P_{LS}(f)$ is the Lomb-Scargle power from \eq{lomb-scargle-periodogram}.
The effect of this exponentiation, as seen in \fig{bayesian-periodogram},
is to suppress all peaks in relation to the largest one of the spectrum.

@@

The problem, however, is that the Bayesian periodogram does {\it not} return
a simple probability that the data oscillates with a given frequency, but rather
a probability of each frequency conditioned on a relatively strong assumption:
that the data are drawn from a sinusoidal model.
Now, the standard Lomb-Scargle periodogram also assumes a sinusoidal model,
and is thus subject to the same criticism.
But unlike the simple Bayesian approach, the standard periodogram affords
interpretation in light of Fourier transforms, window functions, and the
like: in most cases of interest, this turns out to be {\it vital} to the
application and interpretation of the periodogram in most cases of interest
(see \sect{failure-modes}).

In other words, the Bayesian approach goes ``all-in'' on the least squares
interpretation of the periodogram, exponentially suppressing the information
that allows you to reason about the periodogram in light of its connection
to Fourier analysis.
In contrived cases with structureless window functions, exponential attenuation
of side-peaks and side-lobes may seem appealing \citep[see, e.g.][]{Mortier15},
but that appeal extends to the real world only in the most favorable of
cases---{i.e.} high signal-to-noise measurements with a very well-behaved
survey window.

This is not to say that all Bayesian approaches to periodic analysis are
similarly flawed; there have been many interesting studies that go beyond
the simple sinusoidal model and use more complex and/or flexible models.
Some examples are models based on Fourier
extensions with strong priors \citep[e.g.][]{Bretthorst88},
instrument-dependent parametric models \citep[e.g.][]{Angus16},
flexible non-parametric functions \citep[e.g.][]{Gregory92}, and
specially-designed stochastic models \citep[e.g.][]{Kelly14}.
Though these can often be more accurate and powerful than the Lomb-Scargle
approach, they tend to be far more expensive computationally.
Bayesian approaches based on MCMC also tend to run into stability problems,
particularly for multimodal or other complicated posterior distributions
\citep[See, for example, the RR Lyrae discussion in][]{Kelly14}.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig25_failure_modes}
  \caption{Comparison of the true period and peak Lomb-Scargle period for
    1000 simulated periodic light curves.
    Each has 60 irregular observations over 180 nights,
    with a cadence typical of ground-based surveys ({i.e.} showing a strong
    diurnal window pattern similar to \fig{LINEAR-window}).
    The Lomb-Scargle peak does not always coincide with the true period,
    and there is noticeable structure among these failures.
    Panels (a)-(d) isolate some of the specific modes of failure that should
    be expected for this kind of window function;
    see the text for more discussion.
    \figlabel{failure-modes}}
\end{figure}

@@

These patterns can be understood in terms of the interaction between the
window function and the underlying spectral power.
As we discussed in \sect{window-functions},
a nightly observation pattern---typical of ground-based surveys---leads to
a window function with a strong diurnal component that causes each frequency
signature $f_0$ to be partially aliased at $f_0 + n \delta f$, for integers
$n$ and $\delta f = 1\ {\rm day}^{-1}$
(Recall \figs{LINEAR-window}{LINEAR-window-effect}).
This is the source of the failure modes depicted in panel (a) of
\fig{failure-modes}; in terms of periods the above expression becomes:
\begin{equation}
  P_{obs} = \left(\frac{1}{P_{true}} + n\ \delta f\right)^{-1}
\end{equation}
where, to be clear, $n$ is a positive or negative integer and $\delta f$ is
a strong feature in the window function (here, 1 day).
For this data, close to 36\% of the objects are mis-characterized along these
failure modes.

Another mode of failure is when the periodogram for an object with frequency
$f_0$ isolates a higher harmonic of that fundamental frequency, such as $2f_0$.
This may occur for signals that are not strictly sinusoidal, and this harmonic
frequency peak is subject to the same windowing effects as above.
Thus we can describe the failure modes depicted in panel (b) of \fig{failure-modes} with the following equation, for positive integers $m > 0$:
\begin{equation}
  P_{obs} = \left(\frac{m}{P_{true}} + n\delta f\right)^{-1}
\end{equation}
In \fig{failure-modes}, the $m=2$ harmonic and its aliases account for roughly
15\% of the results.

Panel (c) of \fig{failure-modes} shows the final pattern of failure in
Lomb-Scarlge results, which has an opposite trend with true Period.
This is closely related to the effects shown in panels (a) and (b),
but comes from the fact that
the Lomb-Scargle periodogram at frequency $f_0$ contains contributions from
signals at both $+f_0$ and $-f_0$.
When diurnal aliasing leads to power at negative frequencies, this is reflected
at the absolute value of that frequency, so it is more accurate to take the
absolute value of the above formula
\begin{equation}
  P_{obs} = \left|\frac{m}{P_{true}} + n\delta f\right|^{-1}
  \eqlabel{failure-modes}
\end{equation}
Panel (c) shows the 5\% of objects that fall along this failure mode for
$m=1$, $n=-2$.
After accounting for all these known sources of failure, only roughly
1\% of points are misclassified in an ``unexplained'' way, seen in panel (d).

When applying a periodogram in practice, it is vital to take such effects into
account, rather than blindly relying on the single periodogram peak as your
best estimate of the period.
Applying understanding of windowing and aliasing effects can help in detecting
failures of the periodogram, but is no silver bullet.
For an observed peak at $f_{peak}$ from a survey whose window has strong power
at $\delta f$, something like the following should probably be employed:
\begin{enumerate}
  \item Check for a peak at $f_{peak}/m$ for at least $m \in \{2, 3\}$.
    If a significant peak is found, then $f_{peak}$ is probably an order-$m$
    harmonic of the true frequency.
  \item Check for peaks at $|f_{peak} \pm n\delta f|$ for at least
    $n \in \{1, 2\}$. If these aliases exist, then it is likely that you have
    found a peak on the sequence of expected aliases---though keep in mind that
    there is no way to know from the periodogram alone whether or not
    this is the ``true'' peak!
  \item For each of the top few of these aliases, fit a more complicated model
    (such as a multi-term Fourier series, template-based fit, etc.) to select
    among them.
\end{enumerate}
For noisy observations, this procedure cannot generally guarantee that you
have found the correct peak, but it is far preferable than making the mistake
of blindly trusting the highest peak in the periodogram!
We will come back to the question of uncertainty in the periodogram result
in \sect{uncertainties}.

\subsection{Algorithmic Considerations}
\sectlabel{algorithmic-considerations}
Given the size of the frequency grid required to fully characterize the periods
from a given dataset, it is vital to have an efficient algorithm for evaluating
the periodogram. We will see that the na{\"i}ve implementation of
\eq{lomb-scargle-periodogram} scales poorly with the size of the data, but that
fast alternatives are available.

Suppose you have a set of $N$ observations over a time-span $T$ for an average
cadence of $\overline{\delta t} = N/T$.
From \eq{n-eval} we see that the number of frequencies evaluated is
$N_f \propto T f_{max} = N f_{max} / \overline{\delta t}$.
Holding constant the average survey cadence $\overline{\delta t}$ and individual
observation properties (which control $f_{max}$), we find that the number of
frequencies required is directly proportional to the number of data points.
The computation of the Lomb-scargle periodogram in \eq{lomb-scargle-periodogram}
requires sums over $N$ sinusoids at each of the
$N_f$ frequencies, and thus we see that the
na{\"i}ve scaling of the algorithm with the size of the dataset is
$\mathcal{O}(N^2)$, when survey properties are held constant.
Due to the trigonometric functions involved, this turns out to be a rather
``expensive'' $\mathcal{O}(N^2)$, which makes the straightforward
implementation impractical for even modestly-sized datasets.

@@

All periodograms in this paper were computed using the Astropy
implementations of the \citet{Press89}, \citet{Zechmeister09}, and
\citet{Palmer09} algorithms, which were adapted from Python code originally
published by \citet{ICVG2014} and \citet{VanderPlas2015}.
All code behind this paper is available in the form of Jupyter notebooks in the
paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}.

@@

One point of confusion in periodogram implementations is the choice of
normalization.
Recall that there are two equally valid perspectives of what the periodogram
measures: the Fourier view and the least-squares view.
Each of these perspectives lends itself to a different approach to
periodogram normalization.

\subsubsection{PSD normalization}
When considering the periodogram from the Fourier perspective,
it is useful to normalize the periodogram such that in the special case
of equal spaced data, it recovers the standard Fourier power spectrum.
This is the normalization used in \eq{lomb-scargle-periodogram} and the
equivalent least-squares expression seen in \eq{lomb-scargle-chi2}:
\begin{equation}
  P(f) = \frac{N}{2}\big[\hat{\chi}^2_0 - \hat{\chi}^2(f)\big]
  \eqlabel{lomb-scargle-chi2-2}
\end{equation}
For equally-spaced data, this periodogram is identical to the standard
definition based on the fast Fourier transform:
\begin{equation}
  P(f) = \frac{1}{N} \left| FFT(y_n) \right|^2
\end{equation}
in particular, this means that the units of the periodogram are $unit(y)^2$,
and can be interpreted as squared-amplitudes of the Fourier component at
each frequency.
Note that this changes if data uncertainties are incorporated into
the periodogram as in \sect{extensions-observational-noise},
the periodogram in this normalization becomes unitless: it is essentially
a measure of periodic content in signal-to-noise ratio
rather than in signal itself.

\subsubsection{Least-squares normalization}
In the least-squares view of the periodogram, it is more common to normalize
the periodogram so that the values lie between 0 and 1.
When we express the periodogram as a function of $\chi^2$ model fits as in
\eq{lomb-scargle-chi2-2}, it becomes clear that the periodogram has a
mathematical maximum value: if the sinusoidal model {\it perfectly} fits
the data at some frequency $f_0$, then $\hat{\chi}^2(f_0) = 0$ and the
periodogram is maximized at a value of $N\hat{\chi}^2_0 / 2$.
On the other end, it is mathematically impossible for a sinusoidal model to
do worse than the simple constant, non-varying fit reflected in
$\hat{\chi}^2_0$, and so the minimum value of \eq{lomb-scargle-chi2-2} is
exactly zero.

This suggests a different normalization of the periodogram, where the values
are dimensionless and lie between 0 and 1:
\begin{equation}
  P_{norm}(f) = 1 - \frac{\hat{\chi}^2(f)}{\hat{\chi}^2_0}
  \eqlabel{lomb-scargle-normalized}
\end{equation}
This ``normalized periodogram'' is unitless, and
is directly proportional to the unnormalized
(or PSD-normalized) periodogram in \eq{lomb-scargle-chi2-2}.

The main practical consequence of the normalization choice is that the
resulting periodogram has slightly different statistical properties, and
this needs to be taken into account when computing uncertainties and false
alarm probabilities.
Other normalizations exist as well, but seem to be rarely used in practice.
For a summary of these alternative normalizations and their statistical
properties, refer to the introduction of \citet{Baluev2008}.

\subsection{Uncertainties in Periodogram Results}
\sectlabel{uncertainties}
An important piece of reporting results from the Lomb-Scargle periodogram is
the uncertainty of the resulting period estimate.
In most areas of science we are used to being able to report uncertainties in
terms of errorbars, {e.g.} ``the period is $16.3 \pm 0.6$ hours''.
For periods derived from the Lomb-Scargle periodogram, however, uncertainties
generally cannot be meaningfully expressed in this way:
as we saw in the discussion of failure modes in \sect{failure-modes},
the concern for periodograms is more often errors associated with
false peaks and aliases, rather than with imprecision in peak location.

\subsubsection{Peak Width and Frequency Precision}

A periodic signal will be reflected in a periodogram by the presence of a
peak of a certain height and width.
The width of the peak is directly related to the precision with which that
peak's frequency is measured; the height of the peak 
is related to the significance of the detection.
We have motivated from Fourier window discussions that the peak width scales
with the observational baseline; the surprising result is that the peak width
{\it does not depend} on either the number of observations or the noise
in those observations!

This can be proven mathematically \citep[see, e.g.][]{Bretthorst88, ICVG2014},
but it is perhaps more intuitive to see this visually.
\fig{peak-width-height} shows periodograms for simulated data with a fixed
baseline $T=4$ units covering four cycles, with varying sample sizes and
signal-to-noise ratios.
In all cases, the width of the peaks (most clearly visible in one or another
normalization) are essentially {\it identical}, regardless of the data
quality or quantity!
Evidently, peak width ({i.e.} frequency precision) is not a particularly
useful measure of uncertainty in the detection of periods.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig26_peak_width_height}
  \caption{The effect of the number of points $N$ and the signal-to-noise
    ratio $S/N$ on the expected width and height of the periodogram peak.
    Top panels show the normalized periodogram (\eq{lomb-scargle-normalized}),
    while bottom panels show the PSD-normalized periodogram
    (\eq{lomb-scargle-periodogram})
    Somewhat surprisingly, neither the number of points nor the signal-to-noise
    ratio affects the peak width; nor do they affect the peak height under
    the appropriate normalization.
    \figlabel{peak-width-height}}
\end{figure}

If you insist on reporting precision derived from a peak width
in terms of an errorbar, the results can be pretty silly for observations
with long baselines.
For example, going by the peak width in \fig{LINEAR-power}, the
periodogram reveals a period of $2.58024 \pm 0.00005$ hours.
While this is an accurate characterization of the period precision
{\it assuming that peak is the correct one}, it does not capture the
much more relevant uncertainties demonstrated in \sect{failure-modes},
in which we might find ourselves on the wrong peak entirely!
This is why in general, peak widths and Gaussian error bars are not all that
useful for reporting uncertainties in the context of periodogram analysis.

\subsubsection{False Alarm Probability}

A much more relevant quantity for uncertainty in periodogram results is the
height of the peak, and in particular the height compared to the background
noise.
\fig{peak-width-height} indicates that this property {\it does} depend on
both the number of observations and their signal-to-noise ratio: for the
small-$N$/low signal-to-noise cases, the spurious peaks in the background
become comparable to the true peak.
In fact, as we saw in \sect{schuster-to-lomb-scargle}, the ability to
analytically define and quantify the relationship between peak height
and significance is one of the considerations that led to the standard
form of the Lomb-Scargle periodogram.

The typical approach to quantifying the significance of a peak is the {\it
False Alarm Probability} (FAP), which measures the probability that a dataset
with no signal would---due to coincidental alignment among the random
errors---lead to a peak of a similar magnitude.
\citet{Scargle82} showed that for Gaussian noise, the values of the unnormalized
periodogram in \eq{lomb-scargle-periodogram} are exponentially distributed;
that is, for {\it one single frequency} $f_0$, if $Z = P(f_0)$ is the
periodogram value from \eq{lomb-scargle-periodogram}, then
\begin{equation}
  P_{single}(Z) = 1 - \exp(-Z)
  \eqlabel{LS-prob-z}
\end{equation}
is the cumulative probability of observing a periodogram value less than $Z$,
in data with no periodic signal
(be aware that for different periodogram normalizations,
the form of this distribution changes; see \citet{Cumming99} or
\citet{Baluev2008} for a summary for various normalizations).

\subsubsubsection{Independent Frequency Method}

But we are generally not interested in the distribution of one particular
frequency, but rather the distribution of {\it the highest peak} of the
periodogram, which is a quite different situation.
If this is confusing, consider rolling a standard 6-sided die.
The probability, in a single roll, of rolling a number, say $r > 4$ is easy
to compute: it's 2 sides out of six, or $p(r>4) = 1/3$.
If, on the other hand, you roll 10 dice and choose the largest number among
them, the probability that it will be greater than 4 is {\it far} larger
than $1/3$; it is $p(max(r,10) > 4) = 1 - (1 - 1/3)^{10} \approx 0.98$.
The case for the periodogram is analogous: the chance of getting a high peak
at any random location (\eq{LS-prob-z}) is relatively small, but if you search
among a number of large frequencies it is much larger.

With the dice, the calculation is easy because the number of independent
rolls is clear.
With the periodogram, though, the value at one frequency is correlated with
the value at other frequencies in a way that is quite difficult to analytically
quantify---these correlations come from the convolution with the survey
window.
Nonetheless, one common approach to estimating the distribution has been to
assume it can be modeled on some ``effective number'' of independent
frequencies $N_{eff}$, so that the FAP can be estimated as
\begin{equation}
  FAP(z) \approx 1 - \big(P_{single}(z)\big)^{N_{eff}}
  \eqlabel{FAP-neff}
\end{equation}
A very simple estimate for $N_{eff}$ is based on our arguments of the
expected peak width, $\delta f = 1/T$.
In this approximation, the number of independent peaks in a range
$0 \le f \le f_{max}$ is assumed to be $N_{eff} = f_{max} T$.
There have been various attempts to estimate this value more carefully,
both analytically and via simulations
\citep[see, e.g.][]{Horne86,Schwarzenberg-Czerny98,Cumming04,Fresura2008},
but all such approaches are necessarily only approximations.

\subsubsubsection{Baluev (2008) Method}
